On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-08-29 Revisiones/PNA_Responses to the reviewers.tex

\item \textit{The development is very confusing and would make more sense were it to be tied to 
the way that GIMP methods are usually stated, i.e. that the standard FE shape functions are replaced by weighting functions comprised of the conjugation of a standard shape function and a particle "characteristic" function. Line 155 demonstrates this, i.e. the use of $N_{Ip}$ before explanation.}\\

The development of the MPM approach has been revised considering the proposal of the reviewer.

\item \textit{There is little discussion on some of the pitfalls of using max-ent, e.g. the need for a convex domain, the issue of determining Lagrange Multiplier solutions for points close to boundaries, i.e. the solutions process here at Eqns 37 and 38.}\\

Of course, the employment of the Max-Ent shape function is not free of difficulties. Some requirements such as the need for a convex domain, as well as those derived from its calculation (the determination of the Lagrange Multipliers, specially at the boundaries, or the obtainment of the minimizer of the logarithmic function), make the usage of Max-Ent shape functions a challenging tool.

Additionally, Arroyo \& Ortiz~\cite{Arroyo2006} briefly discussed the existence of non-convex domains, and proposed some solutions to it. 
For instance, the possibility of replacing the Euclidean distance $\lVert  x - x_a  \rVert$ in the definition of the shape functions by the length of the shortest path contained within the domain connecting $x$ and $x_a$. Or by decomposing the non-convex domain into convex sub-domains.  On the other hand, this topic has been extensively studied in the context of MLS-based meshfree methods, for instance visibility, diffraction, and constrained path criteria.  These methods are directly applicable to local \textit{max-ent} approximation.

However, the results depicted in this manuscript will strengthen the motivation of its employment in order to mitigate typical MPM problems such as cell-crossing or stress instabilities.

@@

\bibitem{Sadeghirad_2011}
Sadeghirad, A. and Brannon, R. M. and Burghardt, J.
\newblock {A convected particle domain interpolation technique to extend applicability of the material point method for problems involving massive deformations}.
\newblock {International Journal for Numerical Methods in Engineering}, 86(12) : 1435-1456, 2011.

\bibitem{HAMMERQUIST2017724}
Chad C. Hammerquist and John A. Nairn.
\newblock {A new method for material point method particle updates that reduces noise and enhances stability}.
\newblock {\em Computer Methods in Applied Mechanics and Engineering},  318:724--738, 2017.

\bibitem{MA2009272}
S. Ma and X. Zhang and X.M. Qiu.
\newblock{Comparison study of MPM and SPH in modeling hypervelocity impact problems}.
\newblock{International Journal of Impact Engineering}, 36(2):272--282, 2009.

\bibitem{Arroyo2006}
Arroyo, M. and Ortiz, M..
\newblock{Local maximum-entropy approximation schemes: A seamless bridge between finite elements and meshfree methods}.
\newblock{International Journal for Numerical Methods in Engineering},  2006.

### 2020-08-29 LME-MPM.tex
Pdf revisado y hoja de respuestas actualizada. MMP

\usetikzlibrary{%
    decorations.pathreplacing,%
    decorations.pathmorphing%
  }

@@

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method} 
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{cpdi}{CPDI}{Convected Particle Domain Interpolation}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}
\newacronym{dbc}{DBC}{Dirichlet boundary conditions}
\newacronym{nbc}{NBC}{Neumann boundary conditions}

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.

@@

\MODIFIED{Now, employing the particle integral definition \eqref{eq:particle_riemann_integral} to compute the integral of \eqref{eq:particle_acceleration_forces_II}, and simplifying the it, is possible to achieve a discrete expression o the acceleration forces term as}

@@

\subsection{Impact of two elastic discs}
\label{sec:Impact-two-elastic-discs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\sidecaption
 \centering
 \resizebox{0.7\hsize}{!}{
   \input{Figures/Sulsky_discs.tex}}
 \caption{Geometrical description of the two colliding discs problem.}
 \label{fig:Sulsky_discs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\usetikzlibrary{%
    decorations.pathreplacing,%
    decorations.pathmorphing%
  }
\usetikzlibrary{arrows.meta}

@@

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method} 
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{pic}{PIC}{Particle in Cell}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{cpdi}{CPDI}{Convected Particle Domain Interpolation}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{usl}{USL}{Update Stress Last}
\newacronym{usf}{USF}{Update Stress First}
\newacronym{fe}{FE}{Forward Euler}
\newacronym{dbc}{DBC}{Dirichlet boundary conditions}
\newacronym{nbc}{NBC}{Neumann boundary conditions}

@@

The aforementioned techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious,
numerical noises appear despite of the employment of these
techniques~\cite{Tran2019e}. These numerical inaccuracies, also known
as wiggles, arise due to inaccuracies in the time discretisation technique.
A simple approach to face those spurious noises lies on the addition of nonphysical damping sources to the equilibrium equations. This
approach has been widely employed in this and many other numerical
techniques. To avoid introducing this nonphysical sources, many
researchers have proposed alternative time integration
schemes which reduce the presence of high frequency noises by
filtering them or increasing somehow the accuracy. One of the first attempts was
the proposal of an implicit time integration scheme by Guilkey \&
Weiss \cite{Guilkey_2003}. More recently Wang {\it et al.}
\cite{Wang_2016} mitigated these spurious noises by adding a non-viscous
damping to the linear momentum balance equation, and later Charlton
{\it et al.} \cite{Charlton_2017} extended this scheme to the
\acrshort{gimp} approach introducing the \acrfull{igimp}
method. However, the local damping introduced by 
\cite{Wang_2016} can totally over-damp the solution in time-dependent
simulations such as in consolidation process. In a recent publication,
Kularathna \& Soga \cite{Soga_2017} studied an implicit treatment of the pressure in MPM algorithm to simulate material incompressibility
avoiding artificial pressure oscillations by applying
Chorin\textquotesingle s
projection. Within the explicit time integration schemes, Lu {\it et al.}\cite{LU_2018} introduced the time-discontinuous Galerkin method to control the spurious noises
propagation, and later Tran \& Solowski~\cite{Tran2019e}
proposed a generalised-$\alpha$ scheme for \acrshort{mpm} with
promising results but at the expense of increasing the computational
effort. In this paper a less time consuming and high efficient explicit predictor-corrector integration method is
proposed. It consists of an accommodation of the traditional \acrfull{npc} scheme, widely employed in Finite Element methods. This method
has been chosen among other suitable alternatives as those proposed
by Wilson {\it et al.} (1972)~\cite{Wilson1972} or Chung \& Hulbert
(1993)~\cite{Geranlized_alpha_1993} because its simplicity and good
performance dealing with solid dynamic problems under a meshfree
framework in~\cite{Navas2018a}.  
\MODIFIED{ Other well known source of energy dissipation and numerical noise is the presence of a non-trivial null space of the linear operator that maps particles values onto nodal values. It was earlier identified by Brackbill~\cite{BRACKBILL1988469} as ringing instability in the \acrfull{pic} method. A recent development in the field of particle methods is a null-space filter engineered by Gritton \&
Berzins~\cite{Gritton2017} which overcome this limitation but still introduce unwelcome damping.  Later, Hammerquist \& Nairn~\cite{HAMMERQUIST2017724} introduced the XPIC as a parametric extension of the aforementioned research with excellent results.}

@@

\MODIFIED{Now, introducing the particle integral definition \eqref{eq:particle_riemann_integral} to compute the integral of \eqref{eq:particle_acceleration_forces_II}, and simplifying the it, is possible to achieve a discrete expression o the acceleration forces term as}

@@

\MODIFIED{\subsection{Impact of two elastic discs}}
\label{sec:Impact-two-elastic-discs}

\MODIFIED{The following test is best suited to validate the ability of the discussed
interpolation technique to deal with grid crossing instabilities. Also, it enlightens the energy conservation properties of the integration scheme, since it should remains constant before and after collision. The test proposed by Sulsky {\it et al.} (1994) \cite{Sulsky1994} consists in the impact of two identical elastic discs with the following elastic parameters :
\begin{itemize} 
\item  Initial density : $20$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $10^5$
\end{itemize}
}
\MODIFIED{The discs starts in the lower left and upper right corners with initial velocities of $(1, 1)$ and $(-1, -1)$, respectively. The computational domain consists in a square of 10 meters. It is discretised with a cartesian grid with a meshsize of 0.1. The courant number is fixed to 0.1 during the whole simulation. Details of the problem are sketched in figure~\ref{fig:Sulsky_discs}.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
 \centering
 \resizebox{0.7\hsize}{!}{
   \input{Figures/Sulsky_discs.tex}}
 \caption{Geometrical description of the two colliding discs problem.}
 \label{fig:Sulsky_discs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\MODIFIED{Energy plots in figure~\ref{fig:Sulsky-energy} proofs in one hand the ability of the \acrshort{lme} approximation to mitigate the grid-crossing error as well as the \acrshort{ugimp} interpolants, but without some  of its shortcomings previously mentioned in this paper. And in the other hand the energy conservation is also guaranteed with the \acrshort{npc} time integration scheme, unlike what happens in the \acrfull{usf} and \acrfull{usl} schemes. Although during impact mechanical energy decreases, no real dissipation occurs since after collision the value is again 5 Jules like before collision. Furthermore, kinetic energy has been exactly computed. Therefore, this apparently dissipation of mechanical energy is due to a numerical cancellation during the computation of the mechanical energy during post-process stage. }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Sulsky_energy}
  }
  \caption{Energy plots for different interpolation/approximation techniques. All simulations where performed with the already tested \acrshort{npc} algorithm. The evolution of the total energy of the system is represented with the solid line, the dashed line shows the evolution of the deformation energy, while the doted line represents the evolution of the kinetic energy.} 
  \label{fig:Sulsky-energy}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
